---
title: "R Lab I"
author: "Zeren Li"
date: "9/2/2019"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls()) # remove all object
```

## Roadmap

  - R Markdown
  - Seeing theory
  - Exploring CEO salary dataset
  - Problem set
  
## R Markdown

- This is an R Markdown (http://rmarkdown.rstudio.com) Notebook.

- Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. 

- R Markdown mainly consists of three parts: YAML header, texts, and `r` code chunck.

- R Markdown usually begins with a  YAML header (optional) surrounded by `---`s, the header specifies meta information.

- You can write your texts with features like using header `#`, *italic*, **bold** , etc.

- When you run code within [R Markdown](http://rmarkdown.rstudio.com), the results show below the chunk of code.

- You can set chunk global options that apply to every chunk in your file. This is done by calling `knitr::opts_chunk$set` in this code chunk. However, these global defaults can be overwritten in individual chunk headers.

- To understand more chunck options like `echo = TRUE`, `message = FALSE`, and `warning = FALSE`, check [RMarkdown tips and tricks](https://rmarkdown.rstudio.com/lesson-3.html). 

- Insert a new chunk: click the *Insert Chunk* button & using *Cmd+Option+I*.

- Execute chunk: click the *Run* button within the chunk or using *Cmd+Shift+Enter*. 

- Click the **Knit** button to generate a document that includes both contents as well as the output of any embedded R code chunks within the document. 

## Seeing Theory

"**Seeing Theory** is a project designed and created by Daniel Kunin with support from Brown University's Royce Fellowship Program. The goal of the project is to make statistics more accessible to a wider range of students through interactive visualizations."

Check this: https://seeing-theory.brown.edu/basic-probability/index.html

### Importing dataset

Here are various ways of importing data:

```{r read_data, message=FALSE}
# load packages
library(readr) # read csv,txt files
library(tidyverse) # powerful set of data science packages
library(haven) # read stata files
library(stargazer) # regression table
# set working directory (set your own directory)
setwd("./")

# read RData (R)
load("UNpop.RData")

# read csv
UNpop <- read_csv("./UNpop.csv") # readr package

# read dta (Stata)
UNpop_stata_new <- read_dta("UNpop.dta") # haven package (new)
```

###  Read CEO data

```{r}
ceo = read_dta("./CEOSAL2.DTA") # read CEO dataset
```

### View Data

```{r view_data, eval=FALSE}
# View data
View(data)
```

### Explore CEO data

```{r}
class(ceo) # type of object

names(ceo) # variable names (column)

nrow(ceo) # number of rows

ncol(ceo) # number of columns

summary(ceo)  # summarize the dataset

summary(ceo$salary)  # summarize the variable

length(ceo) # length of a dataset means # of variables

length(ceo$salary)  # length of a variable means # of obs

head(ceo) # show the first 5 rows of the dataset

table(ceo$grad) # show the frequency of a categorical variable

ceo_grate <- ceo[ceo$grad == 1,] # fiter by condition(s)

ceo_over_1kk <- ceo[ceo$salary > 1000,] # fiter by conditionn(s)

ceo_1to5 <- ceo[c(1:5), ] # filter by index

ceo_1 <- ceo[,c("salary", "profmarg")] # select by variable name

ceo_var1to5 <- ceo[,c(1:5)] # select by index

# rename variable  
names(ceo_1)
names(ceo_1)[2] <- "profit_margin" 
names(ceo_1)

rm(ceo_1) # remove dataset
```

A bit beautiful summary statistics...

```{r, results =  "asis"}
stargazer(data.frame(ceo))  # summarize the variable
```


### Mean and Variance 

population mean: 

$$\mu = \frac{\sum_{i=1}^n{x_i}}{n}$$ 

sample mean: 

$$\bar x = \frac{\sum_{i=1}^n{x_i}}{n}$$ 


```{r}
m_salary <- sum(ceo$salary)/length(ceo$salary)
m_salary
mean(ceo$salary)
```

population variance:
$$
\sigma^2 = \text{E}[(X - \text{E}[X])^2]
$$

sample variance:

$$
s^2 = \frac{\sum_{i=1}^n(x_i - \bar x)^2}{n-1}
$$

```{r}
sum( (ceo$salary - m_salary)^2 ) / (length(ceo$salary)-1)
var(ceo$salary) # R computes sample variance
```

### Covariance & Correlation

population covariance:

$$Cov(X,Y) = E[(X-E(X)(Y-E(Y))]$$

sample covariance:

$$Cov(X,Y) = \frac{ \sum_{i=1}^n(x_i -\bar x)  (y_i -\bar y)}{n-1} $$

We would like to look at the covariance and correlation between CEO's salary and firm performance measured by profit margins.

```{r}
cov(ceo$salary,ceo$profmarg) # covariance

m_profmarg = sum(ceo$profmarg)/length(ceo$profmarg)
sum((ceo$salary - m_salary) * (ceo$profmarg - m_profmarg ))/(length(ceo$profmarg) - 1) 
```


$$Corr(X,Y) = \frac{E[(X-E(X)E(Y-E(Y))]}{\sqrt{Var(X)Var(Y)}}$$

```{r}
cor(ceo$salary,ceo$profmarg) # correlation

# How to compute manually? 
# Show it in the problem set, it should be the same as the result from cor()
```

# R graph

```{r}
plot(ceo$salary) # one-way scatterplot

plot(ceo$salary, ceo$profmarg) # two-way scatterplot

hist(ceo$salary, main = "Histogram of CEO's salary") # histogram

plot(density(ceo$salary), main = "Density estimate of CEO's salary") # pdf

plot(ecdf(ceo$salary),main = "Empirical cumulative distribution function") # cdf
```

## Other resources

Installing RMarkdown: https://bookdown.org/yihui/rmarkdown/

Frequently asked questions: https://yihui.name/knitr/faq/

RMarkdown cheatsheet: https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf

R Style: http://adv-r.had.co.nz/Style.html