---
title: "Lab-8"
author: "Zeren Li"
date: "10/24/2019"
output: html_document
---

## Roadmap

- Dummy variable
- Categotical variable
- Interaction effect

```{r}
library(wooldridge)
library(tidyverse)
library(ggplot2)
library(viridis)
library(GGally)
library(AER)
```


### Teaching Rating Dataset

This is a dataaset on course evaluations, course characteristics, and professor characteristics for 463 courses for the academic years 2000--2002 at the University of Texas at Austin.


```{r}
data(data = "TeachingRatings", package = "AER")

```



```{r}
TeachingRatings %>%
  select(eval, tenure, gender) %>%
  as.data.frame() %>%
  ggpairs()
```


```{r, results='asis'}
stargazer::stargazer(TeachingRatings, header = F)
```



```{r}
m1 <- lm(eval ~ beauty  , TeachingRatings )

summary(m1)
y_hat <- predict(m1)

ggplot(TeachingRatings, aes(x = beauty, y = eval)) +
  geom_point() +
  geom_smooth(method = "lm" )
```

/*Take a good look at this. Now, the dummy plays a
different role, modifying the impact of TeachingRatings
What  two lines I can draw out of this regression?*/

Dummies are best thought of as shifts in the constant. Thus, 
the new intercept for the female line is the coefficient on catcher plus the original intercept.

$$  eval = \beta_{0} + \beta_{1}beauty + \delta female + e$$

$$  E(eval) = 
    \begin{cases}
      (\beta_{0} + \delta)  + \beta_1 beauty., & \text{female}=1 \\
      \beta_{0} +\beta_1 beauty., & \text{female} = 0
    \end{cases} $$

What about substantive effects. What is our predicted eval for the 
mean female and the mean male from our dummy variable regression?

```{r}
m2 <- lm(eval ~ beauty + gender , TeachingRatings )

# male intercept
m2$coefficients[1] 
# female intercept
m2$coefficients[1] +  m2$coefficients[3]
```


For male with average beauty score : $\beta0 + \beta1*mean(exper)$  

```{r}
 m2$coefficients[1] + m2$coefficients[2] * mean(TeachingRatings$beauty)
```

For female with average beauty score: $\beta0 + \beta1*mean(exper) + \beta2 female$  

```{r}
m2$coefficients[1] + m2$coefficients[2] * mean(TeachingRatings$beauty) + m2$coefficients[3]
```

Notice the different intercepts we predicted, but the slopes are **similar**.

*Why is this method different then running a regression on 
only females and then only males and comparing the results 
with two lines.
pause;

```{r}
TeachingRatings %>% 
  filter(gender == "female")  %>% 
   lm(eval ~ beauty,.)   


TeachingRatings %>% 
  filter(gender == "male")  %>% 
   lm(eval ~ beauty,.)   

# reg eval exper if female==1;
ggplot(TeachingRatings, aes(x = beauty, y = eval, group = gender, color = factor(gender) )) +
  geom_point() +
  geom_smooth(method = "lm" ) 
```


Now, what if I had two dummies in this regression model.  Let's
add another model for above average TeachingRatings. For both male and female

```{r}
lm(eval ~ beauty + gender + tenure, TeachingRatings) %>% summary()
```

For male without tenure (female = 0, tenure = 0)

For male with tenure (female = 0, tenure = 1)

For female without tenure (female = 1, tenure = 0)

For female with tenure (female = 1, tenure = 1)


## Multiple categorical variables

### Create age group

Here, we create an `age_group` categorical variable. If a professor is in 30-39, then we code `age_group` as 1; If 40-49, we code `age_group` as 2, If 50-59, we code `age_group` as 3; If 60 and above, then 4

```{r}
tr1 <- TeachingRatings %>%
  mutate(age_group  = if_else(age <40 ,1 ,
                      if_else(age <50 ,2,
                      if_else(age <60 ,3, 4))) %>% as.factor())

table(tr1$age_group)

```



```{r}
ggplot(tr1, aes(x = age_group, fill = age_group )) +
  geom_bar()
```


```{r}
lm(eval ~ age_group + beauty, tr1) %>% summary()
```

## Interaction Analysis

### Marginal Effect 

Marginal effects tells us how a dependent variable (outcome) changes when a specific independent variable (explanatory variable) changes. Other covariates are assumed to be held constant. Marginal effects are often calculated when analyzing regression analysis results.


### The marginal effect of `beauty` on `eval`?

$$
\frac{\partial \hat{eval}}{\partial beauty} = \hat{\beta}_1
$$

Note: `beauty` is a continuous variable

Answer: One unit increase in `beauty` is associated with $\hat{\beta}_1$  change in `eval`.

### Question 1b

Hypothetically, if `beauty` is a dummy variable (say, `beauty=1` means good-looking and `beauty=0` means otherwise), what is the marginal effect?

Answer: When `beauty` change from 0 to 1, `eval` change by $\hat{\beta}_1$. 



## Interacting dummy variable with dummy variable

$$\hat{eval_i} = \hat{\beta}_0  + \hat{\beta}_1 female_i +  \hat{\beta}_2 tenure_i + \hat{\beta}_3 female_i*tenure_i + u $$

```{r}
# two equaviliant way of interaction
m0 <- lm(eval ~ gender*tenure ,  TeachingRatings)
summary(m0)

m1 = lm(eval ~  gender + tenure + gender:tenure ,  TeachingRatings)
summary(m1)
```

$$\hat{eval_i} = \hat{\beta}_0  + \hat{\beta}_1 female_i +  \hat{\beta}_2 tenure_i + \hat{\beta}_3 female_i*tenure_i + u $$

### Substantive Effects of Dummy Interactions

\begin{table}
\centering
\begin{tabular}{lll}  \hline
 & untenured  &  tenured \\ \hline
male & \hat{\beta}_0 = 4.44 & \hat{\beta}_0 +  \hat{\beta}_2 = 4.44 + (-0.41) = 4.03 \\ 
female &  \hat{\beta}_0  + \hat{\beta}_1  = 4.40 + (-0.54) = 3.86 & & \hat{\beta}_0  + \hat{\beta}_1 \hat{\beta}_2 + \hat{\beta}_3  = 4.4 + (-0.5) + (-0.4) + (0.46) = 3.91 \\ \hline
\end{tabular}
\end{table}

## Interacting dummy variable with continous variable

$$
eval_i = \beta_0 + \beta_1 beauty_i + \beta_2 gender + \beta_3 gender_i \times beauty_i+ u
$$

```{r}
m_int1 <- lm(eval ~  beauty + gender + gender:beauty ,  TeachingRatings)
summary(m_int1)
```

$$
eval_i = \beta_0(4.08595) + \beta_1(0.20027) beauty_i + \beta_2(-0.19510) gender + \\ \beta_3(-0.11266) gender_i \times beauty_i+ u
$$

###  The marginal effect of `beauty` on `eval`

$$
\frac{\partial \hat{eval}}{\partial beauty} = \hat{\beta}_1 + \hat{\beta}_3 \times gender = 0.20027 + -0.11266 = 0.08761
$$

Note: This is a *continuous* variable.

Answer: One unit increase in `beauty` is associated with $\hat{\beta}_3 + \hat{\beta}_5 \times gender$ units increase of `eval`.

### Exercise: the marginal effect of `gender` on `eval`

$$
\frac{\partial \hat{eval}}{\partial gender} = \hat{\beta}_2 + \hat{\beta}_3 \times beauty
$$

Note: This is a *dummy* variable which only takes values 0 and 1.

Answer: When `minority` change from 0 to 1, `eval` change by $\hat{\beta}_4 + \hat{\beta}_5 \times age$.


## Interacting continous variable with continous variable


$$
eval_i = \beta_0 + \beta_1 beauty_i + \beta_2 age + \beta_3 beauty_i\times age_i+ u
$$

###  The marginal effect of `beauty` on `eval`

```{r}
m_int2 <- lm(eval ~  beauty + age + beauty:age ,  TeachingRatings)
summary(m_int2)
```

Our fitted model is:

$$
eval_i = \beta_0(3.9904777) + \beta_1(-0.3391625) beauty_i + \beta_2(0.0006434) age \\
+ \beta_3(0.0101498) beauty_i\times age_i+ u
$$


$$
\frac{\partial \hat{eval}}{\partial beauty} = \hat{\beta}_1 + \hat{\beta}_3 \times age = 
-0.3391625 + 0.0101498 = -0.3290127
$$

### Exercise: What's the marginal effect of age

$$
\frac{\partial \hat{eval}}{\partial age} = ?
$$

## Plotting the Marginal Effect of a Variable

Back to the business: How do we plot the interaction effect?? Let's set it straight, what are the variables of the x-axis and the y-axis?

### Marginal effect of a continuous variable that has no interaction effect

Example: We would like to plot the marginal effect of `beauty` on `eval` using the following model 

$$
eval_i = \beta_0 + \beta_1 beauty_i + \beta_2 age + u
$$

- x-Axis: Value of ?? (anything, the marginal effect is constant conditional on any other independent variables)
- y-Axis: Value of $\frac{\partial \hat{eval}}{\partial beauty} = \beta_1$

No interaction effect: This means the marginal effect of ``beauty` is constant. There isn't much value to visualize it 

```{r}
library(margins)
m <- lm(eval ~ age + beauty + tenure + gender, data = TeachingRatings)
margins(m)
plot(margins(m))
```

### Visualing the marginal effect in an interaction model

For example, we want to visualize this model:

$$
eval_i = \beta_0(3.9904777) + \beta_1(-0.3391625) beauty_i + \beta_2(0.0006434) age \\
+ \beta_3(0.0101498) beauty_i\times age_i+ u
$$

Example: We would like to plot the marginal effect of `beauty` on `eval`

- x-Axis: Value of `age`
- y-Axis: Value of $\hat{\beta}_1 + \hat{\beta}_3 \times age$


```{r}
m_int2 <- lm(eval ~  beauty + age + beauty:age ,  TeachingRatings)

plot(margins(m_int2, at = ))
```

```{r}
# Step 0: Does the marignal effect depend on any other variables?
# Yes. Age

# Step 1: Get summary statistics of the variables identified in the above step
summary(d$age)

# Step 2: Simulate a sequence of the variable of interest in its range
age_sim = seq(min(d$age), max(d$age), length.out=20)

# Step 3: Multiply the sequence with the simulated beta's
quantiles = c(0.025, 0.5, 0.975)

me_minority = apply(beta_sim[, c("age", "age:minority")], 1, function(x) x["age"] * 1 + x["age:minority"] * age_sim) %>%
  apply(1, function(x) quantile(x, quantiles)) %>%
  as.data.frame() %>%
  t() %>%
  as.data.frame() %>%
  mutate(variable = age_sim)

me_minority

me_minority %>% 
  ggplot(aes(x = variable, y = `50%`)) +
  geom_line() + geom_errorbar(aes(ymin = `2.5%`, ymax = `97.5%`), color = "darkgray") +
  geom_hline(yintercept = 0, color = "red", linetype = 2) +
  xlab("age") + ylab("Marginal Effect of Minority Status")

me_minority %>% 
  ggplot(aes(x = variable, y = `50%`)) +
  geom_line() + geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), alpha = 0.3) + 
  geom_hline(yintercept = 0, color = "red", linetype = 2) +
  xlab("age") + ylab("Marginal Effect of Minority Status")
```

### Marginal effect of a continuous variable that interacts with a dummy variable

Example: We would like to plot the marginal effect of `minority` on `eval`

- x-Axis: Value of `beauty`
- y-Axis: Value of $\hat{\beta}_4 + \hat{\beta}_5 \times age$

```{r}
# Step 0: Does the marignal effect depend on any other variables?
# Yes. minority

# Step 1: Get summary statistics of the variables identified in the above step
table(d$minority)

# Step 2: Simulate a sequence of the variable of interest in its range
minority_sim = c(0, 1)

# Step 3: Multiply the sequence with the simulated beta's
quantiles = c(0.025, 0.5, 0.975)

me_age = apply(beta_sim[, c("minority", "age:minority")], 1, function(x) x["minority"] * 1 + x["age:minority"] * minority_sim) %>%
  apply(1, function(x) quantile(x, quantiles)) %>%
  as.data.frame() %>%
  t() %>%
  as.data.frame() %>%
  mutate(variable = minority_sim)

# Plot: using the original code: Bad
# Problem: An informative plot for a variable interacting with a dummy is slightly different.
me_age %>% 
  ggplot(aes(x = variable, y = `50%`)) +
  geom_line() + geom_errorbar(aes(ymin = `2.5%`, ymax = `97.5%`), color = "darkgray") +
  geom_hline(yintercept = 0, color = "red", linetype = 2) +
  xlab("minority") + ylab("Marginal Effect of Age")

me_age %>% 
  mutate(variable = factor(variable)) %>%
  ggplot(aes(x = variable, y = `50%`)) +
  geom_point() +
  geom_line(aes(group = 1)) + geom_errorbar(aes(ymin = `2.5%`, ymax = `97.5%`), width = 0.1) +
  geom_hline(yintercept = 0, color = "red", linetype = 2) +
  xlab("minority") + ylab("Marginal Effect of Age")
# This also apply for interaction with categorical variables (more than two classes)
```
