---
title: "Lab 7"
author: "Zeren Li"
output:
  pdf_document: default
---

```{r, include=FALSE}
library(AER) # For the testsi
library(wooldridge)
library(tidyverse)
library(GGally)
library(stargazer)
library(car)
theme_set(theme_minimal())

knitr::opts_chunk$set(
      echo = TRUE,
      warning=FALSE,
      message=FALSE,
      fig.width=5, 
      fig.height=4 )
```

## Roadmap

- Functional Programming in R
- Goodness of Fit
- Joint Significance Tests

## Function in R

Functions are defined by code with a specific format:

```{r eval=FALSE}
function.name <- function(arg1, arg2, arg3=2, ...) {
  newVar <- sin(arg1) + sin(arg2)  # do some useful stuff
  newVar / arg3   # return value 
}
```


- **function.name** is the function's name. This can be any valid variable name, but you should avoid using names that are used elsewhere in R

- **arg1, arg2, arg3**: objects inside the `()` are the arguments of the function, also called input. You can write a function with any number of arguments. These can be any R object: numbers, strings, arrays, data frames, or even pointers to other functions

- **function body**: The function code between the within the `{}` brackets is run every time the function is called. This code might be very long or very short. Ideally, functions are short and do just one thing – problems are rarely too small to benefit from some abstraction. Sometimes a large function is unavoidable, but usually, these can be in turn constructed from a bunch of small functions.

- **Return value**: The last line of the code is the value that will be returned by the function. A function doesn't need to return anything, for example, a function that makes a plot might not return anything, whereas a function that does a mathematical operation might return a number or a list.

```{r}
# a simple sum of square function
sum_of_squares <- function(x,y) {
  x^2 + y^2
}

x1 <- 3
y1 <- 4

data("saving" , package = "wooldridge")
  
# specify the return value
sum_of_squares1 <- function(x,y) {
  x_2 <- x^2; y_2 <- y^2
  z <- x_2 + y_2
  return(c(z))
}

sum_of_squares1(x1,y1)
```

What makes for a good function? READABLE

- Make it Short
- Perform a single operation
- Use intuitive names

### Exercise: write your own bivariate OLS function 

```{r}
ols_bh <- function(x,y) {
  b1 = mean(cov(x,y)/var(x)) %>% round(5)
  b0 = mean(y - b1*x) %>% round(5)
  return(paste("beta1 is", b1, ", and beta0 is", b0 ))
}

x <- rnorm(10,10,2)
y <- rnorm(10,5,1)

ols_bh(x,y)

lm(y ~ x)
```

## Goodness of Fit

### Root Mean Square Error (RMSE) 

Root Mean Square Error is the standard deviation of the residuals (prediction errors). This metric is commonly used for prediction analysis.

\begin{align*}
RMSE = \sqrt{ \frac 1 {n- k - 1}  \sum_i{(y_i - \hat y)^2}}
\end{align*}

- $k$: the number of regressors *excluding* the intercept.
- $n$: the number of observations
- $\hat y$: the predicted value of y

### Adjusted R^2

- $R^2$ increasing whenever an additional regressor is added to the model
    - Adding a regressor ---> $SSR$  goes down ---> $R^2$ goes up
- Adjusted $R^2$ "punishes" the addition of regressors using a correction factor, which is defined as

$$ \bar{R}^2 = 1-\frac{n-1(1-R^2)}{n-k-1}  = 1-\frac{n-1}{n-k-1} \, \frac{SSR}{TSS}. $$

##  Evaluation on Test Set

* Split the data into 2 groups:  training and test

* Fit model to training data

* Predict on test data

* Evaluate RMSE on test data

* In OLS, training RMSE can be driven smaller by adding more predictors

* How well does model predict for new data $y^*$?

\begin{align*}
RMSE^* = \sqrt{ \frac 1 {n^*- k^* - 1}  \sum_i{(y^*_i - \widehat{y})^2}}
\end{align*}

* No guarantee that model that has the lowest training RMSE will be the best out of sample RMSE. 

* If test RMSE is much larger than training RMSE, the model might suffer from overfitting.

## MLB

```{r}
data("mlb1", package="wooldridge")  #from wooldridge package

# variable definition
?mlb1

mlb <- mlb1 %>%
  # add population variable
  mutate(total_pop=(100*whitepop)/ percwhte)

```

1. Set a random number seed at any positive integer to start the randomization process. This will allow me to repeat the process
and always obtain the same numbers

```{r}
n = nrow(mlb)
n.train = floor(.75*n) # 75% training set, 25% test set
set.seed(2019) 
train = sample(1:n, size=n.train, rep=F)

mlb_train <- mlb[train,]
mlb_test <- mlb[-train,]
```

### Exploratory data analysis

```{r}
mlb_train %>% 
    select("years", "gamesyr", "bavg", "hrunsyr", "rbisyr")   %>% 
ggpairs()
```

### Summary statistics

```{r, results= "asis"}
mlb_train %>% 
    select("years", "gamesyr", "bavg", "hrunsyr", "rbisyr")   %>%
    stargazer(., header = F, title = "Summary Statistics")
```

## Fit a model

- The baseline model of player salary. This is where we will begin my analysis.  

```{r}
m1 <- lm(lsalary ~ years + gamesyr, mlb_train )
yhat <- predict(m1)

# compute rmse
rmse_bh <- function(y, ypred,k) {
          rmse = sqrt(sum((y - ypred)^2)/(length(y)-k-1))
          return(rmse)
}

rmse_bh(mlb_train$lsalary, yhat,2)
summary(m1)
```

- Then, we add my key theoretical causal variables. Batting Average,
Home Runs, and RBIs, all in the most recent year. 

```{r}
m2 <- lm(lsalary ~ years + gamesyr + bavg + hrunsyr + rbisyr , mlb_train )
summary(m2)
```


# Compute adjusted R^2 by hand

```{r}
# define the components
n <- nrow(mlb_train)                            # number of observations (rows)
k <- length(m2$coefficients)-1                  # number of regressors
y_mean <- mean(mlb_train$lsalary)

SSR <- sum(residuals(m2)^2)               # sum of squared residuals
TSS <- sum((mlb_train$lsalary -  y_mean)^2)       # total sum of squares
ESS <- sum((fitted(m2) - y_mean)^2)       # explained sum of squares
# compute the measures

Rsq <- 1 - (SSR / TSS)                          # R^2
adj_Rsq <- 1 - (n-1)/(n-k-1) * SSR/TSS          # adj. R^2

# print the measures to the console
c("R2" = Rsq, "Adj.R2" = adj_Rsq)

# results using built-in function
summary(m2)$r.squared
summary(m2)$adj.r.squared

# Can we write an entire function that gives us r squared, adjusted r squared, and rmse?
```

## Joint hypothesis testing using the F-Statistic

- Can we reject the hypothesis that the coefficients are zero? 

- A joint hypothesis imposes restrictions on multiple regression coefficients. 

- This is different from conducting individual $t$-tests where a restriction is imposed on a single coefficient. 

The homoskedasticity-only $F$-Statistic:

$$ F = \frac{(SSR_{\text{restricted}} - SSR_{\text{unrestricted}})/q}{SSR_{\text{unrestricted}} / (n-k-1)} $$;

$$ F ~ F_{q,n-k-1}$$ 

- $SSR_{restricted}$ being the sum of squared residuals from the restricted regression, i.e., the regression where we impose the restriction. 
- $SSR_{unrestricted}$ is the sum of squared residuals from the full model
- $q$ is the number of restrictions 
- $k$ is the number of regressors in the unrestricted regression

```{r}
# bulit-in function
linearHypothesis(m2, c("bavg=0", "hrunsyr=0", "rbisyr=0"))

# compute by hand
SSR_r <- sum(residuals(m1)^2)              
SSR_ur <- sum(residuals(m2)^2)               
q <- 3      # number of restriction
n <- nrow(mlb_train)            # number of observations (rows)
k <- length(m2$coefficients)-1  # number of regressors
(F_stats <- ((SSR_r - SSR_ur) / q)/(SSR_ur / (n - k - 1)))

# get p value
1 - pf(F_stats, df1 = q, df2 = n - k - 1)

```

Why are the baseball stats insignificant in individual t-testing, even
though they are jointly significant?

Our gut reaction should be that they are highly collinear. Those who 
hit more home runs are more likely to have more runs batted in

Indeed, they are.  Look at the VIF on RBIs.  It is almost entirely 
explained by the other variables.

Indeed, home runs and rbis have a .89 correlation.

```{r}
vif(m2)

cor(mlb_train$hrunsyr, mlb_train$rbisyr)
```

Theoretically, we choose to drop rbis, because rbis are a function of other players, whereas home runs are only created by a single player. What happens to the coefficient and standard errors on home runs after I do this?

```{r}
m3 <- lm(lsalary ~ years + gamesyr + bavg + hrunsyr, mlb_train)
summary(m3)
```

How about adding some other measures of baseball success? e.g. runs scored per year, stolen bases per year, and career fielding perc.

```{r}
m4 <- lm(lsalary ~ years + gamesyr + bavg + hrunsyr + runsyr + fldperc + sbasesyr, mlb_train)
summary(m4)
```

None of the three are individually significant at the .05 level, and the variable of stolen bases has the wrong sign? Why would someone who steals a lot of bases be paid less?

```{r}
vif(m4)
cor(mlb_train$sbasesyr, mlb_train$runsyr)
```

Are fielding percentages and stolen bases doing anything for us?
Let's do an F-test of the joint significance of these variables.

```{r}
m5 <- lm(lsalary ~ years + gamesyr + bavg + hrunsyr + runsyr + fldperc + sbasesyr , mlb_train)

linearHypothesis(m5, c("fldperc=0","sbasesyr=0"))
```

We cannot reject the null hypothesis at the .05 level.

The sign flip on sbases is probably due to high collinearity with runs (lead-off hitters tend to have high runs and stolen bases) and fielding percentage. We choose to drop fldperc and sbasesyr.

So far we haven't yet controlled for the race of the player or the market size, 
might these have an impact? 

These other factors are important, what sort of bias
might they be causing in my regressions. Think this through carefully.

```{r}
m6 <- lm(lsalary ~ years + bavg + hrunsyr + runsyr + total_pop + pcinc + hispan + black, mlb_train)
summary(m6)

linearHypothesis(m6, c("total_pop=0","pcinc=0","hispan=0","black=0"))
```

No, they are jointly insignificant. Demographic factors do not seem to affect salary.

How about a position. Short-stop is a highly skilled position. What if we tested whether short-stops are more highly paid than other positions?

```{r}
m7 <- lm(lsalary ~ years + gamesyr + bavg + hrunsyr + runsyr + shrtstop, mlb_train)
summary(m7)
linearHypothesis(m7, c("shrtstop=0"))

# How about catcher?
m8 <- lm(lsalary ~ years + gamesyr + bavg + hrunsyr + runsyr + catcher, mlb_train)

summary(m8)
# Variance Inflation Factors
vif(m8)

rmse_bh(m8$model$lsalary, predict(m8),6)
summary(m8)
linearHypothesis(m8, c("catcher=0"))
```

Yes, all else equal, catchers are paid about 40% more than all players.

What does all else equal or the oft-used *ceteris paribus* mean?   It means that for players with similar years and games played and equivalent offense statistics, catchers get paid more.  **It does not mean just being a catcher means you get paid more.**


### Prediction using test data

```{r}
# yhat of training data
y_train_hat <- predict(m8)
rmse_train <- rmse_bh(mlb_train$lsalary, y_train_hat,6)

# yhat of test data
y_test_hat <- predict(m8, newdata  = mlb_test)

rmse_test <- rmse_bh(mlb_test$lsalary, y_test_hat,6)

c(rmse_train = rmse_train , rmse_test = rmse_test)
```

## Take-way

\centering \large \textbf ``All models are wrong, but some are useful'' -- George Box

